---
title: "Blog 2: Programming Background"
author: "Michael Bradshaw"
date: "`r Sys.Date()`"
---

```{r setup, include=TRUE}
knitr::opts_chunk$set(fig.path = "C://Users//mbradshaw//OneDrive - Research Triangle Institute//Documents//Graduate School Courses//ST 558//Blog Update//mikebrad140.github.io//images/")
```
```{r createblog, eval=FALSE, echo=FALSE}
rmarkdown::render(
  input = "C://Users//mbradshaw//OneDrive - Research Triangle Institute//Documents//Graduate School Courses//ST 558//Blog Update//mikebrad140.github.io//_RMD//2023-06-06-blog-post-number2.Rmd",
  output_format = "github_document",
  output_dir = "C://Users//mbradshaw//OneDrive - Research Triangle Institute//Documents//Graduate School Courses//ST 558//Blog Update//mikebrad140.github.io//_posts",
  output_options = list(keep_html = FALSE, html_preview=FALSE)
)
```  
## Thoughts on the R programming language.  

As someone who has worked primarily as a SAS programmer for the past 15 years, I'm very excited to add R to my data science skillset. Although SAS is widely used within my company and industry, R is another language that is widely used. My experience with R has been very positive. R feels very powerful and user friendly. The intergrated help system within R, and the vast amount of books and communities that help programmers learn really helps speed up the learning curve.

I'm excited to learn more and continue to practice learning how to write my own functions and also learn how to manipulate lists. This feels like the missing link for me with getting proficient in R. 

Compared to SAS, I feel like the ggplot and other data visualization options are far superior. R feels more flexible and allows for more customization over plot aesthetics. Additionally, the integration of R Markdown allows for creating reports that combine text, code, and graphical output for dissemination. This is highly applicable as a public health data analyst.

Currently, I'm still getting used to thinking about data manipulation in R. My brain is wired towards SAS, where you use data steps, procedures, and/or SQL to manipulate datasets. Additionally, I tend to think about how I can macrotize my code, versus in R, where it seems you need to use functions, lists, and loops to do this.

R has a wealth of learning resources available, including online tutorials, books, forums, and communities. This extensive documentation can assist in learning and troubleshooting.

I'm really enjoying learning R and getting comfortable processing and manipulating data. R does seems like there is a learning curve. It's not necessarily hard, but just getting comfortable with the object oriented programming apporach. I think the hardest part for me is going to be taking the leap and starting to program in R when I know it will take longer compared to SAS. However, I think this is the only way for me to truly learn this language.

## Example R Markdown output.

You can also embed plots, for example:

```{r iris, echo=FALSE}
plot(iris)
```

