---
title: "Project 1 Review Blog:"
author: "Michael Bradshaw"
date: "`r Sys.Date()`" 
---

```{r setup, eval=TRUE, echo=FALSE}
knitr::opts_chunk$set(fig.path = "../images/")
```

```{r createblog, eval=FALSE, echo=FALSE}
rmarkdown::render(
  input = "C://Users//mbradshaw//OneDrive - Research Triangle Institute//Documents//Graduate School Courses//ST 558//Blog Update//mikebrad140.github.io//_RMD//2023-06-23-blog-post-number-Project1.Rmd",
  output_format = "github_document",
  output_dir = "C://Users//mbradshaw//OneDrive - Research Triangle Institute//Documents//Graduate School Courses//ST 558//Blog Update//mikebrad140.github.io//_posts",
  output_options = list(keep_html = FALSE, html_preview=FALSE)
)
```  
#### Explaining the project and interesting findings:  

This project involved contacting the Spoonacular API to pull recipe information based on our search criteria. Our functions allowed users to search recipes by query(keyword), cuisine, diet, ingredients we want to include, ingredients we want to exclude, how long the recipe takes to make, and the maximum number of recipes we want returned. The Spoonacular API's extensive database of recipes and nutrition information allowed us to pull in detailed information about these recipes, like taste information, and nutrition information. The search queries allowed users to receive recipes that catered to individual culinary tastes and dietary restrictions.

After contacting different parts of the Spoonacular API and returning the parsed data based on our search criteria, we combined our returned data and created a dataset that allowed us to perform exploratory data analysis on our recipes. First, we created a couple of new variables including a made-up "Yummy Score" and a Yummy Categorical variable based on the recipes "Yummy Score". Then we looked at summary tables for different variables to better understand differences and similarities in our search results among key variables. Finally, we created five plots to further examine distributions and relationships between our recipe characteristics. 

One interesting finding was looking at relationships between variables based on their Yummy Category.  I was surprised that super yummy recipes had lower fat content that the other yummy categories. Additionally, I thought it interesting that there were such few vegetarian options. I suppose that excluding tomatoes from our search criteria could have contributed to this lack of vegetarian options in our returned query.


#### Reflecting on the process:
The process of producing the code and analysis was a valuable learning experience, especially working with an API. Learning to interact with an API in R via a PowerPoint presentation is one thing, but actually applying that knowledge in practice often presents different challenges. I learned the importance of understanding the API documentation and learning about the available endpoints. Additionally, this process of applying API specific knowledge in practice requires trial and error, problem-solving, and adapting to specific API nuances (like search limits), which was not discussed in the our notes.

#### Most difficult part of the logic and programming for you?  

The most challenging part for me was integrating the Spoonacular API effectively into the project. Understanding the API documentation, retrieving and parsing the required data, and implementing the necessary functionality required a lot of browsing the API documentation resources and a bit of trial and error. Additionally, each query of the Spoonacular API costs points and the API sets a cap on the number of points available to free users per day. I was also not familiar with how to effectively use the unlist() command to take lists and create a single variable from a long list. For example, cuisine and diet both returned lists from the API. 

#### What would I do differently in approaching a similar project in the future?  

There seems to be a wide variety in the available API documentation online. Before getting into the data, I think it would be important to spend more time understanding the structure of the API and think how best to integrate the returned data into a suitable dataset for analysis. This would help me plan the project more effectively and avoid potential roadblocks or inefficiencies during implementation.  

Finally, I would have explored additional ways to personalize the recipe search queries even further. Expanding the scope of personalization would make the system even more valuable to users with diverse needs and dietary preferences.  

#### Final thoughts: 
Through this project, I gained valuable knowledge and experience, expanding my understanding of API integration in R and honing my problem-solving skills. I am better equipped to tackle similar projects more effectively in the future.  

[Link to my ST558 Project 1 github pages repo](https://github.com/mikebrad140/ST558---Project-1)  

[Link to my ST558 Project 1 non-github pages site](https://mikebrad140.github.io/ST558---Project-1/)