---
title: "Project 2 Review Blog:"
author: "Michael Bradshaw"
date: "`r Sys.Date()`" 
---

```{r setup, eval=TRUE, echo=FALSE}
knitr::opts_chunk$set(fig.path = "../images/")
```

```{r createblog, eval=FALSE, echo=FALSE}
rmarkdown::render(
  input = "C://Users//mbradshaw//OneDrive - Research Triangle Institute//Documents//Graduate School Courses//ST 558//Blog Update//mikebrad140.github.io//_RMD//2023-07-07-blog-post-number-Project2.Rmd",
  output_format = "github_document",
  output_dir = "C://Users//mbradshaw//OneDrive - Research Triangle Institute//Documents//Graduate School Courses//ST 558//Blog Update//mikebrad140.github.io//_posts",
  output_options = list(keep_html = FALSE, html_preview=FALSE)
)
```  

### What would you do differently?

I think I was too focused about automating the code and should have waiting longer to finish all the modeling before proceeding with automation, as the instructions specified.

### What was the most difficult part for you?
The automation part was tricky for me. The notes were helpful, but making sure the render code linked to the .RMD file was the part I struggled with. In particular making sure the paramaters were correct.It took me some time to troubleshoot and make sure everything was working as intended

Additionally, I kept thinking I was doing something wrong becuase I was expecting the models to do a better job at modeling the shares variable. It took me awhile to convince myself that the predictors are simply fairly weak at predicting shares. I thought the tree models would be significantly better than they were. 

### What are your big take-aways from this project?

The automation of markdown reports in the biggest thing I will take away from this project. The ability to generate reports automatically will prove invaluable in future projects and save considerable time and effort.

Additionally, I learned that the predictive performance of different models can vary, but in this particular case, they were all relatively similar in predicting the shares variable. This was a bit disappointing, but it highlights the importance of thoroughly evaluating and selecting the appropriate models for a given task.

[Link to my ST558 Project 2 github pages repo](https://github.com/mikebrad140/ST558---Project-2)  

[Link to my ST558 Project 2 non-github pages site](https://mikebrad140.github.io/ST558---Project-2/)
